---
title: "hwk1"
author: "Esther Koffi"
date: "2025-02-03"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
q1 <- matrix(c(4, 4, 1, 8, 7, 5, 8, 8, 5, 9, 8, 6), nrow = 6, ncol = 2, byrow = TRUE)
dist(q1)


```

b)
The closest three observations are 3, 4, and 5 

c)
Yes

```{r}
vgsales <- read.csv(file = "vgsales.csv")


```
a-

```{r}
dim(vgsales)

table(vgsales$Genre)

```

16593 observations and 2346 sports games.

b-

```{r}
summary(vgsales$NA_Sales)


```


```{r}
var(vgsales$NA_Sales)

```

c-
```{r}
hist(log(vgsales$NA_Sales), main = "Histroram of natrual logarithm of North American sales")

```

d-
```{r}

pairs(vgsales [ ,5:7])

```

In general,  relationship look positive between the sales for the vgsales data set.

Question 3-


```{r}

bp <- read.csv(file = "BP.csv")
bp$race <- as.factor(bp$race)

```

a-

```{r}
fit0 <- lm(bp ~ age + height + weight + race, data = bp)
summary(fit0)




```

bp estimated = 25.679 - 0.043 * age + 1.312 * height + 0.059 * weight - 2.628 * race3 -4.539 * race4 -3.466* race5


b)
Mean blood pressure increases by 1.312 for each additional inch of height with other variables held constant in this relationship

c)
Height and weight.

d)Adjusted R square = 0.3605 therefore we can said that  the model explains 36.05% of the variance in blood pressure.

e-
```{r}
newdat <- data.frame(age = 20, height = 70, weight = 160, race = as.factor(4))
predict(fit0, newdat)


```

f-
```{r}

fit1 <- lm(bp ~ age + height + weight + race + age:weight, data = bp)
summary(fit1)

```

the adjusted r squared  increase slightly and the interaction is significant at 0.1

g-

```{r}
fit2 <- lm(bp ~ . + log(age) + I(weight^2), data = bp)
summary(fit2)


```
Including all of the variables plus the logarithm of age and a quadratic term in weight increased Adjusted R squared to 0.394. However, log(age), weight^2 and several other predictors were not significant.